---
title: Bootstrap example
output:
    html_document:
        code_folding: show
---

```{r knitr_options, include=FALSE}
knitr::opts_chunk$set(fig.width=11, fig.height=6.5,
                      message=FALSE, warning=FALSE)
options(width=110)
library(broman) # for myround()
set.seed(20200324)
```


This is a short example to illustrate the bootstrap. We'll use the
example from [lecture](https://kbroman.org/AdvData/17_bootstrap.pdf),
of estimating the 95th percentile of a population.

I'll first simulate some data from a scaled chi-squared distribution.
I simulate data from a chi-square with 5 degrees of freedom and
multiply it by 6. I'll simulate a sample of size _n_=800.

```{r simulate_data}
n <- 800
df <- 5
mult <- 6
dat <- rchisq(n, df=df)*mult
```

Here's a histogram of the data.

```{r histogram_data}
par(mar=c(3.1, 1.1, 1.1, 1.1))
hist(dat, breaks=2*sqrt(n), xlab="", ylab="", yaxt="n", main="")
```

The 95th percentile is `r myround(quantile(dat, 0.95), 1)`. (Note that
the true percentile for the underlying population is
`r myround(qchisq(0.95, df)*mult, 1)`.)


### Non-parametric bootstrap

Let's start with the more generic non-parametric bootstrap.
We'll perform 1000 bootstrap replicates.

I'll use `sample()` to sample with replacement from the data. I can do
all 1000 bootstrap replicates at the same time, and then take the 95th
percentile of each column.

```{r nonparametric_bootstrap}
n_boot <- 1000
boot_data <- matrix( sample(dat, n_boot*n, replace=TRUE), ncol=n_boot )
boot_results <- apply(boot_data, 2, quantile, 0.95)
```

Here's a histogram of the bootstrap results. The tick marks underneath
the histogram are the locations of my actual data points. Gaps between
data points are what leads to the gaps in the histogram.

```{r bootstrap_histogram}
par(mar=c(3.1, 1.1, 1.1, 1.1))
hist(boot_results, breaks=2*sqrt(n_boot), main="", ylab="", yaxt="n",
     xlab="estimated 95th percentile")
rug(dat)
```

The SD of these results is `r myround(sd(boot_results), 2)`, so that's
my estimate of the standard error of my estimate.


### Parametric bootstrap

Suppose I used were to model the data as a scaled chi-square
distribution. How could I use the parametric bootstrap in that case?

Well I first need to estimate the two parameters to get a fitted
model: the degrees of freedom and the multiplier. I could use the
method-of-moments for this.

The variance of the scaled chi-squared is (2 m^2^ df) and the mean is
(m df). So I could estimate m by var/mean/2 =
`r myround(m_hat <- var(dat)/mean(dat)/2, 2)` and then estimate df by
mean/m = `r myround(df_hat <- mean(dat)/m_hat, 2)`.

I then simulate from the fitted model with those estimates as the
parameters, and then take the quantile of simulated data set.

```{r parametric_bootstrap}
pboot_data <- matrix( rchisq(n_boot*n, df_hat)*m_hat, ncol=n_boot )
pboot_results <- apply(pboot_data, 2, quantile, 0.95)
```

Here's a histogram of the bootstrap results. The tick marks underneath
the histogram are the locations of my actual data points, but note
that we don't see the odd gaps that we'd seen before.

```{r parametric_bootstrap_histogram}
par(mar=c(3.1, 1.1, 1.1, 1.1))
hist(pboot_results, breaks=2*sqrt(n_boot), main="", ylab="", yaxt="n",
     xlab="estimated 95th percentile")
rug(dat)
```

The SD of these results is `r myround(sd(pboot_results), 2)`, so that's
my estimate of the standard error of my estimate, by the parametric
bootstrap.

### Session details

Below are the versions of R and R packages I used.

The R Markdown source for this document
is [on GitHub](https://github.com/kbroman/AdvData/blob/master/17_bootstrap/17_bootstrap_example.Rmd).

```{r session_info}
devtools::session_info()
```
