\documentclass[aspectratio=169,12pt,t]{beamer}
\usepackage{graphicx}
\setbeameroption{hide notes}
\setbeamertemplate{note page}[plain]
\usepackage{listings}

\input{../LaTeX/header.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% end of header
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Wrangling messy data files}
\author{\href{https://kbroman.org}{Karl Broman}}
\institute{Biostatistics \& Medical Informatics, UW{\textendash}Madison}
\date{\href{https://kbroman.org}{\tt \scriptsize \color{foreground} kbroman.org}
\\[-4pt]
\href{https://github.com/kbroman}{\tt \scriptsize \color{foreground} github.com/kbroman}
\\[-4pt]
\href{https://twitter.com/kwbroman}{\tt \scriptsize \color{foreground} @kwbroman}
\\[-4pt]
{\scriptsize Course web: \href{https://kbroman.org/AdvData}{\tt kbroman.org/AdvData}}
}

\begin{document}

{
\setbeamertemplate{footline}{} % no page number here
\frame{
  \titlepage

\note{
  In this lecture, we'll look at the problem of wrangling messy data
  files: A bit of data diagnostics, but mostly how to reorganize data
  files.
}
} }


\begin{frame}[c]{}

\centering
\Large

{
\color{title}

``In what form would you like the data?''
}

\bigskip \bigskip

\onslide<2->{
{\hilit

    ``In its present form!''
}
}

\bigskip \bigskip \bigskip \bigskip
\onslide<3>{
\lolit
\large
\qquad \qquad \qquad   ...so we'll have some messy files to deal with.
}

\note{
  When collaborators ask me how I would like them to send me data, I
  always say: in its present form.

  I cannot emphasize enough: If any transformation needs to be
  done, or if anything needs to be fixed, it is the data scientist who
  is in the best position to do the work, reproducibly and without
  introducing further errors.

  But that means I spend a lot of time mucking about with some pretty
  messy files.  In the lecture today, I want to impart some tips on
  things I've learned, doing so.
}

\end{frame}


\begin{frame}{Challenges}

  \bigskip

  {\large \vhilit Consistency}

  \bigskip

  \bi
\item file names
\item file organization
\item subject IDs
\item variable names
\item categorical data
  \ei

    \note{
      Essentially all of the challenges come from inconsistencies: in
      file names, the arrangement of data within files, the subject
      identifiers, the variable names, and the categories in
      categorical data.

      Code re-organizing data is the worst code.
    }

\end{frame}





\end{document}
